A Closed-Form Solution to the Doppler Imaging Problem

### 2021-09-29 src/ms.tex
luhman16b

% Author list
\author[0000-0002-0296-3826]{Rodrigo Luger}
\email{rluger@flatironinstitute.org}
\affil{Center~for~Computational~Astrophysics, Flatiron~Institute, New~York, NY}
%
\author{Megan Bedell}
\affil{Center~for~Computational~Astrophysics, Flatiron~Institute, New~York, NY}
%
\author{Daniel Foreman-Mackey}
\affil{Center~for~Computational~Astrophysics, Flatiron~Institute, New~York, NY}
%
\author{Lily Zhao}
\affil{Center~for~Computational~Astrophysics, Flatiron~Institute, New~York, NY}
%
\author{David W. Hogg}
\affil{Center~for~Computational~Astrophysics, Flatiron~Institute, New~York, NY}

@@

Our input spectrum $\bvec{s}$ was generated at high resolution ($R \equiv \frac{\lambda}{\Delta\lambda} = 430,000$) centered on the Fe I $\lambda 6430$ absorption line (blue curve in the bottom panel of Figure~\ref{fig:spot_setup}).
%
The spectrum consists of a (purely synthetic) Gaussian absorption line of fractional depth 50\% centered at $\lambda = 643$~nm with FWHM of $0.02$~nm, or, in velocity space, $\Delta v = \mathrm{FWHM} \times c / \lambda \approx~10$~km/s, equal to the value used in \citet{Vogt1987}. 
%
Unlike \citet{Vogt1987}, we add two additional small absorption lines, one of which is blended with the central absorption line.
In total, the rest frame spectrum $\bvec{s}$ consists of $K = 200$ wavelength bins in the range $642.85~\mathrm{nm} \leq \lambda \leq 643.15~\mathrm{nm}$ plus an additional
%
\begin{align}
    W & = 2 \bigg\lceil \frac{(v\sin i)_\mathrm{max}R}{c} \bigg\rceil \nonumber  \\
      & = 144
\end{align}
%
bins of padding (72 on either side) to eliminate edge effects due to the convolution, for a total of $K' = 344$ wavelength bins. 
In the expression above, $(v\sin i)_\mathrm{max}$ is the maximum value of $v\sin i$ we can model with a given value of $W$; for this experiment, we set $(v\sin i)_\mathrm{max}=50$~km/s, which is larger than the actual value of $v\sin i$ we used to generate the data (see above).
Finally, we linearly interpolate this spectrum to a wavelength grid uniform in log space to obtain the vector $\bvec{s}$ evaluated at each $\lnlam$.

@@

\subsection{Dependence on inclination}
\label{sec:inc}
%
\begin{figure}[t!]
    \begin{centering}
        \includegraphics[width=0.98\linewidth]{figures/inclinations.pdf} % 
        \caption{%
            Recovered \texttt{SPOT} map as a function of stellar inclination. 
            The observed spectra are continuum-normalized and the rest frame spectrum is assumed to be known exactly.
            In general, it is very difficult to infer features in the hemisphere facing away from the observer (in this case, the southern hemisphere). At low inclinations, those features are never in view, while at inclinations close to 90$^\circ$, it becomes difficult to differentiate between features above and below the equator.
        }
        \label{fig:inclinations}
    \end{centering}
\end{figure}
%
As in \citet{Vogt1987}, we also investigate the dependence of the solution on the inclination of the star.
Figure~\ref{fig:inclinations} shows the inferred map for the same problem setup as before, but this time varying the inclination between $10^\circ$ (nearly face-on) and $90^\circ$ (perfectly edge-on). 
We also add four small circular features at $-45^\circ$ latitude to test our ability to infer features in the southern hemisphere (i.e., the side of the star facing away from the observer).
We set the covariance of the spatial prior to $\boldsymbol{\Lambda}_\bvec{y} = 3\times 10^{-5}\bvec{I}$ and, as before, assume the inclination to be known exactly. 
We further keep the true equatorial velocity of the star constant at $60$~km/s across all panels, so that the figure corresponds to the same star viewed at different orientations. 
Because of this, decreasing the inclination decreases $v\sin i$, reducing the width of the convolution kernel and therefore the spatial resolution of our inferred map.
Increasing the inclination does not result in appreciable changes until an inclination of $\sim 80^\circ$, above which it becomes difficult to differentiate between features in the northern and southern hemispheres, whose projected velocities are nearly equal. 
At $i = 90^\circ$, there is a formal degeneracy between the two hemispheres, which can be seen from the mirror-image symmetry in the inferred map. 
Finally, as expected, the features in the southern hemisphere are not recovered below inclinations of about $60^\circ$, as they are never actually in view. At an inclination of $80^\circ$m tge spots are recovered fairly well, but never with as high fidelity as features in the observer-facing hemisphere.
These results are broadly consistent with those of \citet{Vogt1987}.

>>>>

% Author list
\author[0000-0002-0296-3826]{Rodrigo Luger}
\email{rluger@flatironinstitute.org}
\affil{Center~for~Computational~Astrophysics, Flatiron~Institute, New~York, NY}
%
\author[0000-0001-9907-7742]{Megan Bedell}
\affil{Center~for~Computational~Astrophysics, Flatiron~Institute, New~York, NY}
%
\author[0000-0002-9328-5652]{Daniel Foreman-Mackey}
\affil{Center~for~Computational~Astrophysics, Flatiron~Institute, New~York, NY}
%
\author[0000-0002-1835-1891]{Ian Crossfield}
\affil{Department~of~Physics~and~Astronomy, Kansas University, Lawrence, KS}
%
\author[0000-0002-3852-3590]{Lily Zhao}
\affil{Center~for~Computational~Astrophysics, Flatiron~Institute, New~York, NY}
%
\author[0000-0003-2866-9403]{David W. Hogg}
\affil{Center~for~Computational~Astrophysics, Flatiron~Institute, New~York, NY}

@@

Our input spectrum $\bvec{s}$ was generated at high resolution ($R \equiv \frac{\lambda}{\Delta\lambda} = 150,000$) centered on the Fe I $\lambda 6430$ absorption line (blue curve in the bottom panel of Figure~\ref{fig:spot_setup}).
%
The spectrum consists of a (purely synthetic) Gaussian absorption line of fractional depth 50\% centered at $\lambda = 643$~nm with FWHM of $0.02$~nm, or, in velocity space, $\Delta v = \mathrm{FWHM} \times c / \lambda \approx~10$~km/s, equal to the value used in \citet{Vogt1987}. 
%
Unlike \citet{Vogt1987}, we add two additional small absorption lines, one of which is blended with the central absorption line.
In total, the rest frame spectrum $\bvec{s}$ consists of $K = 200$ wavelength bins in the range $642.85~\mathrm{nm} \leq \lambda \leq 643.15~\mathrm{nm}$ plus an additional
%
\begin{align}
    W & = 2 \bigg\lceil \frac{(v\sin i)_\mathrm{max}R}{c} \bigg\rceil \nonumber  \\
      & = 144
\end{align}
%
bins of padding (72 on either side) to eliminate edge effects due to the convolution, for a total of $K' = 344$ wavelength bins. 
In the expression above, $(v\sin i)_\mathrm{max}$ is the maximum value of $v\sin i$ we can model with a given value of $W$; for this experiment, we set $(v\sin i)_\mathrm{max}=50$~km/s, which is larger than the actual value of $v\sin i$ we used to generate the data (see above).
Finally, we linearly interpolate this spectrum to a wavelength grid uniform in log space to obtain the vector $\bvec{s}$ evaluated at each $\lnlam$.

@@

\subsection{Dependence on inclination}
\label{sec:inc}
%
\begin{figure}[t!]
    \begin{centering}
        \includegraphics[width=0.98\linewidth]{figures/inclinations.pdf} % 
        \caption{%
            Recovered \texttt{SPOT} map as a function of stellar inclination. 
            The observed spectra are continuum-normalized and the rest frame spectrum is assumed to be known exactly.
            In general, it is very difficult to infer features in the hemisphere facing away from the observer (in this case, the southern hemisphere). At low inclinations, those features are never in view, while at inclinations close to 90$^\circ$, it becomes difficult to differentiate between features above and below the equator.
        }
        \label{fig:inclinations}
    \end{centering}
\end{figure}
%
As in \citet{Vogt1987}, we also investigate the dependence of the solution on the inclination of the star.
Figure~\ref{fig:inclinations} shows the inferred map for the same problem setup as before, but this time varying the inclination between $10^\circ$ (nearly face-on) and $90^\circ$ (perfectly edge-on). 
We also add four small circular features at $-45^\circ$ latitude to test our ability to infer features in the southern hemisphere (i.e., the side of the star facing away from the observer).
We set the covariance of the spatial prior to $\boldsymbol{\Lambda}_\bvec{y} = 3\times 10^{-5}\bvec{I}$ and, as before, assume the inclination to be known exactly. 
We further keep the true equatorial velocity of the star constant at $60$~km/s across all panels, so that the figure corresponds to the same star viewed at different orientations. 
Because of this, decreasing the inclination decreases $v\sin i$, reducing the width of the convolution kernel and therefore the spatial resolution of our inferred map.
Increasing the inclination does not result in appreciable changes until an inclination of $\sim 80^\circ$, above which it becomes difficult to differentiate between features in the northern and southern hemispheres, whose projected velocities are nearly equal. 
At $i = 90^\circ$, there is a formal degeneracy between the two hemispheres, which can be seen from the mirror-image symmetry in the inferred map. 
Finally, as expected, the features in the southern hemisphere are not recovered below inclinations of about $60^\circ$, as they are never actually in view. 
At an inclination of $80^\circ$ the spots are recovered fairly well, but never with as high fidelity as features in the observer-facing hemisphere.
These results are broadly consistent with those of \citet{Vogt1987}.

\section{Application to Luhman 16B}
\label{sec:luhman16b}
%
\begin{figure}[t!]
    \begin{centering}
        \includegraphics[width=0.45\linewidth]{figures/luhman16b_map.pdf} % 
        \includegraphics[width=0.42\linewidth]{static/luhman16b_crossfield.png} % 
        \caption{%
            Maximum \emph{a posteriori} surface maps of Luhman 16B inferred from the CRIRES spectra obtained in \citet{Crossfield2014}. \emph{Left:} surface map inferred using the algorithm developed in this work. \emph{Right}: surface map inferred in \citet{Crossfield2014}, reproduced from Figure~2 in that paper. We infer the presence of a bright circumpolar spot, as well as the presence of a few dark features, most of which agree with the map presented in \citet{Crossfield2014}.
        }
        \label{fig:luhman16b_maps}
    \end{centering}
\end{figure}
%
